
---
title: "STAT2170/STAT6180 Assignment 2025"
author: "Akib Bin Anowar - 48682462"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

## (a) Correlation and Plot

```{r}
sleep <- read.csv("sleep.csv")
pairs(sleep, main="Pairwise Plots of Sleep Data")
round(cor(sleep), 2)
```

We observe moderate correlation between neck size and arousal index, as well as some positive correlation with age and systolic blood pressure.

---

## (b) Full Regression Model & Neck Size Impact

```{r}
model_full <- lm(ai ~ age + bmi + neck_size + sbp, data=sleep)
summary(model_full)
confint(model_full, "neck_size", level = 0.95)
```

Each cm increase in neck size is associated with an estimated 0.041 increase in log-scaled arousal index (CI: [0.012, 0.069]).

---

## (c) ANOVA and Overall Model Test

Model:  
$ai = \beta_0 + \beta_1 \cdot age + \beta_2 \cdot bmi + \beta_3 \cdot neck\_size + \beta_4 \cdot sbp + \varepsilon$

Hypotheses:  
$H_0$: All $\beta$ = 0 (no relationship)  
$H_1$: At least one $\beta \ne 0$

```{r}
anova(model_full)
```

The F-statistic follows $F_{4,117}$ under $H_0$ and the p-value indicates strong evidence against $H_0$. Thus, at least one predictor significantly explains variation in ai.

---

## (d) Model Diagnostics

```{r}
par(mfrow=c(2,2))
plot(model_full)
```

The residual plots suggest assumptions of linearity, normality, and homoscedasticity are reasonably met.

---

## (e) R-squared

```{r}
summary(model_full)$r.squared
```

About 24.7% of the variation in arousal index is explained by the predictors.

---

## (f) Model Selection

```{r}
step_model <- step(model_full)
summary(step_model)
```

Final model: `ai ~ age + neck_size + sbp`

---

## (g) Model Comparison

The adjusted R² of the reduced model is nearly identical to the full model, suggesting the excluded predictor (bmi) adds little value.

---

# Question 2

## (a) Balanced Design Check

```{r}
energy <- read.csv("energy.csv")
table(energy$range, energy$menu)
```

Each cell has equal observations → the design is **balanced**.

---

## (b) Graphs

```{r}
library(ggplot2)
ggplot(energy, aes(x=factor(range), y=consumption, fill=factor(menu))) +
    geom_boxplot(position=position_dodge()) +
    labs(title="Consumption by Range and Menu")

interaction.plot(x.factor=energy$range, trace.factor=energy$menu,
                 response=energy$consumption,
                 fun=mean, type="b", col=c("red","blue"),
                 legend=TRUE, xlab="Range", ylab="Consumption")
```

---

## (c) Interaction Model

$consumption = \mu + \alpha_i + \beta_j + (\alpha\beta)_{ij} + \varepsilon_{ijk}$  
Where $\alpha_i$ = range effect, $\beta_j$ = menu effect, and $(\alpha\beta)_{ij}$ = interaction.

---

## (d) ANOVA

```{r}
interaction_model <- lm(consumption ~ factor(range)*factor(menu), data=energy)
anova(interaction_model)
```

Significant interaction found (p < 0.001). Model diagnostics are below.

```{r}
par(mfrow=c(2,2))
plot(interaction_model)
```

Conclusion: Both range, menu, and their interaction significantly affect energy consumption.
